---
title: "Guadua Leaf Micromorphology Graphics"
author: "Elizabeth McMurchie"
date: "2022-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load appropriate libraries, installing if needed
```{r}
library(tidyverse) #includes ggplot2 needed for graphics
library(readxl) #to read in doc
library(ade4) #needed for PcoA
library(vegan) #needed for PCoA
library(ggcorrplot) #for correlation tables
```

## Read in data as dataframe
```{r}
guadualeaf <- read_csv("../data/tidy/guadualeaf.csv", col_names = TRUE, na = "x")
```
## Drop columns with NA and no variation and create dataframe with columsn as factors
```{r}
guadualeaf <- guadualeaf %>%
  select(-c(ad_stomata_freq, ad_papillae_overarch, ad_triangular_sub_cells,
            ad_dome_sub_cells, ad_parallel_sub_cells, ab_ridged_saddle_intercostal,
            ad_ridged_saddle_intercostal)) # remove columns with NA or no variation

df_guadualeaf = data.frame(guadualeaf, stringsAsFactors = TRUE) # create dataframe

df_guadualeaf <- data.frame(lapply(df_guadualeaf,as.factor)) # set columns as factors

```

### Correlation
The correlation is necessary to complete the PCoA. Additionally, viewing the correlation dataframe is useful for understanding which traits are correlated.
```{r}
guadua_binary <- df_guadualeaf[,10:ncol(df_guadualeaf)] # reads in the binary data only 
guadua_binary_df <- data.frame(lapply(guadua_binary,function(x) as.numeric(levels(x))[x]))
correlation_guadua <- data.frame(cor(guadua_binary_df)) #calculated the correlation of the Y values

view(correlation_guadua) #viewing basic correlation table

#ggcorrplot alows us to view correlation table with colors indicating correlation values

ggcorrplot(correlation_guadua, hc.order = TRUE, lab = TRUE)

corrs <- ggcorrplot(correlation_guadua, hc.order = TRUE, lab = TRUE)

ggsave("leafcorrplot.pdf", plot = corrs, width = 16, height = 14, units = "in", dpi = 300)

corrs2 <- ggcorrplot(correlation_guadua, hc.order = TRUE, lab = TRUE, type = "upper")

ggsave("leafcorrplot2.pdf", plot = corrs2, width = 16, height = 14, units = "in", dpi = 300)
```

### PCoA 
The distance matrix for the binary data is obtained using simple matching coefficient. The PCoA from the distance matrix can be used to visualize the pairwise distances between the micromorphology of individual *Guadua* specimens.
```{r}
guadua.dist <- dist.binary(guadua_binary_df, method = 2, diag = FALSE, upper = FALSE) # convert to distances using simple matching coefficient
guadua.dist.matrix <- as.matrix(guadua.dist) # create matrix using simple matching distances
PCoA <- cmdscale(guadua.dist.matrix, eig = TRUE, x.ret = TRUE, list. = TRUE) #from vegan - run PcoA
```

# PCoA grouped by habit (3 categories)
```{r}
group.colors <- c("big_erect" = "#7b3294", "leaning_climbing" = "#1b7837",
                  "small_arching" = "#f1a340") #to set colors

nice_df <- PCoA$points %>% 
  data.frame
nice_df$habit <- df_guadualeaf$habit
habitatpcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = habit), size = 2) +
  scale_color_manual(name = "Habit", breaks = c("big_erect", "leaning_climbing", "small_arching"),
                     labels = c("Big erect", "Leaning/climbing", "Small arching"),
                     values = group.colors) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    color = "Habit", 
    x = "Axis 1",
    y = "Axis 2"
  )

ggsave("habits3.pdf", plot = habitatpcoa, width = 6, height = 4, units = "in", dpi = 300)
```

# PCoA grouped by habit ( categories)
```{r}
group.colors2 <- c("erect_or_climbing" = "#7b3294","small_arching" = "#f1a340")

nice_df <- PCoA$points %>% 
  data.frame
nice_df$general_habit <- df_guadualeaf$general_habit
habits2pcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = general_habit), size = 2) +
  scale_color_manual(name = "Habit", breaks = c("erect_or_climbing", "small_arching"),
                     labels = c("Erect or climbing", "Small arching"),
                     values = group.colors2) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    color = "Habit", 
    x = "Axis 1",
    y = "Axis 2"
  )

ggsave("habits2.pdf", plot = habits2pcoa, width = 6, height = 4, units = "in", dpi = 300)

```

# PCoA grouped by habitat (3 categories)
```{r}
group.colors3 <- c("forest" = "#7b3294", "river" = "#1b7837",
                   "savanna" = "#f1a340")

nice_df <- PCoA$points %>% 
  data.frame
nice_df$habitat <- df_guadualeaf$habitat
habitatspcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = habitat), size = 2) +
  scale_color_manual(name = "Habitat", breaks = c("forest", "river", "savanna"),
                     labels = c("Forest", "River", "Savanna"),
                     values = group.colors3) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    color = "Habitat", 
    x = "Axis 1",
    y = "Axis 2"
  )

ggsave("habitat3.pdf", plot = habitatspcoa, width = 6, height = 4, units = "in", dpi = 300)

```

# PCoA grouped by habitat (2 categories)
```{r}
group.colors4 <- c("general_forest" = "#7b3294","savanna" = "#f1a340")

nice_df <- PCoA$points %>% 
  data.frame
nice_df$general_habitat <- df_guadualeaf$general_habitat
habitats2pcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = general_habitat), size = 2) +
  scale_color_manual(name = "Habitat", breaks = c("general_forest", "savanna"),
                     labels = c("Forest or River", "Savanna"),
                     values = group.colors4) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    color = "Habitat", 
    x = "Axis 1",
    y = "Axis 2"
  )


ggsave("habitat2.pdf", plot = habitats2pcoa, width = 6, height = 4, units = "in", dpi = 300)
```

# PCoA grouped by region
```{r}
group.colors5 <- c("andes" = "#7b3294", "central_america" = "#1b7837",
                   "eastern_south_america" = "#af8dc3", "mexico" = "#f1a340")

nice_df <- PCoA$points %>% 
  data.frame
nice_df$region <- df_guadualeaf$region
regionpcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = region), size = 2) +
  scale_color_manual(name = "Region", breaks = c("andes", "central_america",
                                                 "eastern_south_america", "mexico"),
                     labels = c("Andes", "Central America", "Eastern South America", "Mexico"),
                     values = group.colors5) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    color = "Region", 
    x = "Axis 1",
    y = "Axis 2"
  )

ggsave("region.pdf", plot = regionpcoa, width = 6, height = 4, units = "in", dpi = 300)
```

### Scree Plot From the PCoA 
From this we can obtain eigenvalues and then plot them to show the percent variation explained along each PCoA axis. As seen in the plot below, PCoA axis 1 is responsible for the majority of variation in the PCoA, followed by a steep drop in variation. By the 20th axis, variation is arguably negligible. Elbow is at approximately the fourth or fifth principal coordinate axis.
```{r}
eig_df <- data.frame( x_values = c(1:length(PCoA$eig)) , eig_value = c(PCoA$eig))
scree <- ggplot(data = eig_df, aes(x = x_values, y = eig_value)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  labs(
    x = "Coordinate number",
    y = "Eigenvalue"
  )

ggsave("scree.pdf", plot = scree, width = 6, height = 4, units = "in", dpi = 300)
```

