---
title: "Guadua Floret Micromorphology Graphics"
author: "Elizabeth McMurchie and Devin Molnau"
date: "2025-03-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load appropriate libraries, installing if needed
```{r}
library(tidyverse) #includes ggplot2 needed for graphics
library(ade4) #needed for PcoA
library(vegan) #needed for PCoA
library(ggcorrplot) #for correlation tables
```

## Read in data as dataframe
```{r}
guaduafloret <- read_csv("../data/tidy/guaduafloret.csv", col_names = TRUE, na = "x")
```

## Drop columns with NA and no variation and create dataframe with columns as factors
```{r}
guaduafloret <- guaduafloret %>%
  select(-c(lemma_stomata, lemma_stomata_freq, palea_sulcus_macro, palea_stomata_freq, palea_subsidiary_triangular, palea_subsidiary_dome, palea_subsidiary_parallel)) # remove columns with NA or no variation

df_guaduafloret = data.frame(guaduafloret, stringsAsFactors = TRUE) # create dataframe

df_guaduafloret <- data.frame(lapply(df_guaduafloret,as.factor)) # set columns as factors

```

### Correlation
Viewing the correlation dataframe is useful for understanding which traits are correlated. The two-sided correlation plot was used as Appendix C Figure 2 (Figure C.2) in the manuscript. The one-sided correlation plot was not used in the manuscript; this image was used to more easily visualize each correlation.
```{r}
guadua_floret_y_binary <- df_guaduafloret[,10:ncol(df_guaduafloret)] # reads in the binary data only 
guadua_floret_y_binary_df <- data.frame(lapply(guadua_floret_y_binary,function(x) as.numeric(levels(x))[x]))
correlation_y_guadua_floret <- data.frame(cor(guadua_floret_y_binary_df)) #calculated the correlation of the Y values

#view(correlation_y_guadua_floret) #viewing basic correlation table

#ggcorrplot alows us to view correlation table with colors indicating correlation values

#Two-sided correlation plot

ggcorrplot(correlation_y_guadua_floret, hc.order = TRUE, lab = TRUE)

floretcorrs <- ggcorrplot(correlation_y_guadua_floret, hc.order = TRUE, lab = TRUE, ggtheme = theme_bw())

ggsave("floretcorrplot.pdf", plot = floretcorrs, width = 20, height = 18, units = "in", dpi = 600)


ggsave("floretcorrplot.tiff", plot = floretcorrs, width = 20, height = 18, units = "in", dpi = 600)

#One-sided correlation plot

floretcorrs2 <- ggcorrplot(correlation_y_guadua_floret, hc.order = TRUE, lab = TRUE, type = "upper", ggtheme = theme_bw())

ggsave("floretcorrplot2.pdf", plot = floretcorrs2, width = 20, height = 18, units = "in", dpi = 600)

ggsave("floretcorrplot2.tiff", plot = floretcorrs2, width = 20, height = 18, units = "in", dpi = 600)

```

### PCoA 
The distance matrix for the binary data is obtained using simple matching coefficient. The PCoA from the distance matrix can be used to visualize the pairwise distances between the micromorphology of individual *Guadua* specimens.
```{r}
guadua_floret_y_dist <- dist.binary(guadua_floret_y_binary_df, method = 2, diag = FALSE, upper = FALSE) # convert to distances using simple matching coefficient
guadua_floret_y_dist_matrix <- as.matrix(guadua_floret_y_dist) # create matrix using simple matching distances
floret_PCoA <- cmdscale(guadua_floret_y_dist_matrix, eig = TRUE, x.ret = TRUE, list. = TRUE) #from vegan - run PcoA
```

# PCoA grouped by habit (3 categories)
For this PCoA, individual specimens are colored by the habit typical of their species. Three habitat groups were used: big and erect, leaning/climbing, and small arching. This PCoA is used as Figure 10 in the manuscript.
```{r}
group.colors <- c("big_erect" = "#7b3294", "leaning_climbing" = "#1b7837",
                  "small_arching" = "#f1a340") #to set colors

nice_df <- floret_PCoA$points %>% 
  data.frame
nice_df$habit <- df_guaduafloret$habit
habitatpcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = habit, shape = habit), size = 2) +
  scale_color_manual(name = "Habit", breaks = c("big_erect", "leaning_climbing", "small_arching"),
                     labels = c("Tall and erect", "Leaning/climbing", "Small arching"),
                     values = group.colors) + 
  scale_shape_manual(name = "Habit", breaks = c("big_erect", "leaning_climbing", "small_arching"),
                     labels = c("Tall and erect", "Leaning/climbing", "Small arching"),
                     values = c(15, 17, 19)) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) +
  theme(plot.background = element_rect(fill = 'white', color = 'white')) +
  labs(
    color = "Habit", 
    x = "PCoA1",
    y = "PCoA2"
  )

#Remove '#' from below to save images.

ggsave("florethabits3.pdf", plot = habitatpcoa, width = 6, height = 4, units = "in", dpi = 600)

ggsave("florethabits3.tiff", plot = habitatpcoa, width = 6, height = 4, units = "in", dpi = 600)
```

# PCoA grouped by habitat (3 categories)
For this PCoA, individual specimens are colored by the habitat typical of their species. Three habitat groups were used: Forest, river, and savanna. This PCoA is used as Figure 9 in the manuscript.
```{r}
group.colors3 <- c("forest" = "#7b3294", "river" = "#1b7837",
                   "savanna" = "#f1a340")

nice_df <- floret_PCoA$points %>% 
  data.frame
nice_df$habitat <- df_guaduafloret$habitat
habitatspcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = habitat, shape = habitat), size = 2) +
  scale_color_manual(name = "Habitat", breaks = c("forest", "river", "savanna"),
                     labels = c("Forest", "River", "Savanna"),
                     values = group.colors3) + 
  scale_shape_manual(name = "Habitat", breaks = c("forest", "river", "savanna"),
                     labels = c("Forest", "River", "Savanna"),
                     values = c(15, 17, 19)) +
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) +
  theme(plot.background = element_rect(fill = 'white', color = 'white')) +
  labs(
    color = "Habitat", 
    x = "PCoA1",
    y = "PCoA2"
  )

#Remove '#' from below to save images.

ggsave("florethabitat3.pdf", plot = habitatspcoa, width = 6, height = 4, units = "in", dpi = 600)

ggsave("florethabitat3.tiff", plot = habitatspcoa, width = 6, height = 4, units = "in", dpi = 600)

```

# PCoA grouped by species
For this PCoA, individual specimens are colored by species. This PCoA is used as Figure 11 in the manuscript.
```{r}
group.colors <- c("g_aculeata" = "#7b3294", "g_amplexifolia" = "#1b7837",
                  "g_angustifolia" = "#f1a340", "g_calderoniana" = "#00441b","g_chacoensis" = "#5aae61",
                  "g_ciliata" = "#40004b", "g_glomerata" = "#9970ab", 
                  "g_incana" = "#c7e9b4", "g_latifolia" = "#7fcdbb", 
                  "g_longifolia" = "#bd0026", "g_macclurei" = "#fed976", "g_macrospiculata" = "#bfd3e6", 
                  "g_macrostachya" = "#fd8d3c", "g_paniculata" = "#e7298a", 
                  "g_paraguayana" = "#8c96c6", "g_refracta" = "#081d58", "g_sarcocarpa" = "#810f7c", 
                  "g_superba" = "#41b6c4", "g_tagoara" = "#800026", "g_trinii" = "#dd3497",
                  "g_uncinata" = "#1d91c0", "g_velutina" = "#016c59", 
                  "g_venezuelae" = "#feb24c", "g_weberbaueri" = "#78c679", 
                  "g_aff_glomerata" = "#993404",
                  "g_aff_weberbaueri" = "#d4b9da") #to set colors

nice_df <- floret_PCoA$points %>% 
  data.frame
nice_df$species <- df_guaduafloret$species
speciespcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = species, shape = species), size = 2) +
 scale_color_manual(name = "Species", 
                     breaks = c("g_aculeata", "g_amplexifolia","g_angustifolia",
                                "g_calderoniana","g_chacoensis",
                                "g_ciliata", "g_glomerata","g_incana",
                                "g_latifolia", "g_longifolia", "g_macclurei", "g_macrospiculata",
                                "g_macrostachya", "g_paniculata", "g_paraguayana",
                                "g_refracta", "g_sarcocarpa", "g_superba", "g_tagoara", "g_trinii",
                                "g_uncinata", "g_velutina", "g_venezuelae",
                                "g_weberbaueri", "g_aff_glomerata",
                                "g_aff_weberbaueri"),
                                labels = c("G. aculeata", "G. amplexifolia", "G. angustifolia",
                                           "G. calderoniana", "G. chacoensis",
                                           "G. ciliata","G. glomerata", "G. incana",
                                           "G. latifolia", "G. longifolia", "G. macclurei",
                                           "G. macrospiculata", "G. macrostachya","G. paniculata",
                                           "G. paraguayana", "G. refracta","G. sarcocarpa", "G. superba",
                                           "G. tagoara", "G. trinii", "G. uncinata",
                                           "G. velutina","G. venezuelae", "G. weberbaueri",
                                           "G. aff. glomerata",
                                           "G. aff. weberbaueri"),
                     values = group.colors) + 
    scale_shape_manual(name = "Species", 
                     breaks = c("g_aculeata", "g_amplexifolia","g_angustifolia",
                                "g_calderoniana","g_chacoensis",
                                "g_ciliata", "g_glomerata", "g_incana",
                                "g_latifolia", "g_longifolia", "g_macclurei", "g_macrospiculata",
                                "g_macrostachya", "g_paniculata", "g_paraguayana",
                                "g_refracta", "g_sarcocarpa", "g_superba", "g_tagoara", "g_trinii",
                                "g_uncinata", "g_velutina", "g_venezuelae",
                                "g_weberbaueri", "g_aff_glomerata",
                                "g_aff_weberbaueri"),
                                labels = c("G. aculeata", "G. amplexifolia", "G. angustifolia",
                                           "G. calderoniana", "G. chacoensis",
                                           "G. ciliata","G. glomerata", "G. incana",
                                           "G. latifolia", "G. longifolia", "G. macclurei",
                                           "G. macrospiculata", "G. macrostachya","G. paniculata",
                                           "G. paraguayana", "G. refracta","G. sarcocarpa", "G. superba",
                                           "G. tagoara", "G. trinii", "G. uncinata",
                                           "G. velutina","G. venezuelae", "G. weberbaueri",
                                           "G. aff. glomerata",
                                           "G. aff. weberbaueri"),
                     values = c(15, 17, 19, 18, 3, 4, 7, 15, 3, 25, 23, 24, 21, 22, 18, 7, 9, 8, 
                                15, 17, 18, 15, 17, 23, 24, 8)) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) +
  theme(plot.background = element_rect(fill = 'white', color = 'white')) +
  labs(
    color = "Species", 
    x = "PCoA1",
    y = "PCoA2"
  )

plot(speciespcoa)

#Remove '#' from below to save images.

ggsave("floretspecies.pdf", plot = speciespcoa, width = 10, height = 6, units = "in", dpi = 600)

ggsave("floretspecies.tiff", plot = speciespcoa, width = 10, height = 6, units = "in", dpi = 600)

```


# PCoA grouped by region
For this PCoA, individual specimens are colored by general region in which the specimen was collected. Four general regions were used: Andes, Central America, and Eastern South America. Assignment of region to each specimen can be seen in 'specimen_data.xls' in the 'documents' folder. This PCoA was not used in the manuscript. We did not expect to see any grouping based on region, as each region contains a variety of species of different habits and habitats.
```{r}
group.colors5 <- c("andes" = "#7b3294", "central_america" = "#1b7837",
                   "eastern_south_america" = "#af8dc3", "mexico" = "#f1a340")

nice_df <- floret_PCoA$points %>% 
  data.frame
nice_df$region <- df_guaduafloret$region
regionpcoa <- nice_df %>% 
  ggplot(aes(X1, X2)) +
  geom_point(aes(color = region, shape = region), size = 2) +
  scale_color_manual(name = "Region", breaks = c("andes", "central_america",
                                                 "eastern_south_america", "mexico"),
                     labels = c("Andes", "Central America", "Eastern South America", "Mexico"),
                     values = group.colors5) + 
  scale_shape_manual(name = "Region", breaks = c("andes", "central_america",
                                                 "eastern_south_america", "mexico"),
                     labels = c("Andes", "Central America", "Eastern South America", "Mexico"),
                     values = c(15, 17, 19, 18)) + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) +
  theme(plot.background = element_rect(fill = 'white', color = 'white')) +
  labs(
    color = "Region", 
    x = "PCoA1",
    y = "PCoA2"
  )

#Remove '#' from below to save images.

ggsave("floretregion.pdf", plot = regionpcoa, width = 6, height = 4, units = "in", dpi = 600)

ggsave("floretregion.tiff", plot = regionpcoa, width = 6, height = 4, units = "in", dpi = 600)
```

### Scree Plot From the PCoA 
From this we can obtain eigenvalues and then plot them to show the percent variation explained along each PCoA axis. As seen in the plot below, PCoA axes 1 and 2 are responsible for less than half of variation in the PCoA. Elbow is at approximately the eighth principal coordinate axis. Variation becomes negligible at approximately the 30th principal coordinate axis. This image was used as Figure C.4 - Appendix C Figure 4 in the manuscript.
```{r}
eig_df <- data.frame( x_values = c(1:length(floret_PCoA$eig)) , eig_value = c(floret_PCoA$eig))
scree <- ggplot(data = eig_df, aes(x = x_values, y = eig_value)) +
  geom_line() +
  geom_point() + 
  theme_minimal() +
  theme(
    panel.border = element_rect(size = 2, color = "black", fill = NA)
  ) + 
  theme(plot.background = element_rect(fill = 'white', color = 'white')) +
  labs(
    x = "Coordinate number",
    y = "Eigenvalue"
  )

ggsave("floretscree.pdf", plot = scree, width = 6, height = 4, units = "in", dpi = 1000)

ggsave("floretscree.tiff", plot = scree, width = 6, height = 4, units = "in", dpi = 1500)
```

# Pie Charts
The following pie charts show percentage of specimens belonging to different habitat groups for which a certain micromorphological feature was present or absent on foliage leaves. These pie charts were used as components of Figure 5 in the manuscript. 

## Pie charts - specimens from river species

##Filtering dataframe for river species only
```{r}
#Filter dataframe of foliar micromorphology to include only river species
guaduafloret_river <- df_guaduafloret %>%
  filter(habitat == "river")
```

##River: Pie chart of simple papillae on abaxial lemma surface presence/absence
```{r}
#Group by simple papillae on abaxial lemma surface only and obtain percentage present/absent
guaduafloret_river_lemma_papillae_simple_per <- guaduafloret_river %>% 
  group_by(lemma_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_river_lemma_papillae_simple_per)

#Create labeled pie chart of percentage of river specimens with simple papillae on lemma abaxial surface
ggplot(guaduafloret_river_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on lemma abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


river_pie_lemma_simple_papillae <- ggplot(guaduafloret_river_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on lemma abaxial surface")) +
  ggtitle("River") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("river_pie_lemma_simple_papillae.tiff", plot = river_pie_lemma_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##River: Pie chart of simple papillae on abaxial palea surface presence/absence
```{r}
#Group by simple papillae on abaxial palea surface only and obtain percentage present/absent
guaduafloret_river_palea_papillae_simple_per <- guaduafloret_river %>% 
  group_by(palea_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_river_palea_papillae_simple_per)

#Create labeled pie chart of percentage of river specimens with simple papillae on palea abaxial surface
ggplot(guaduafloret_river_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


river_pie_palea_simple_papillae <- ggplot(guaduafloret_river_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on palea abaxial surface")) +
  ggtitle("River") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("river_pie_palea_simple_papillae.tiff", plot = river_pie_palea_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##River: Pie chart of bilobate horizontally elongated silica bodies on abaxial palea surface presence/absence
```{r}
#Group by bilobate horizontally elongated silica bodies on abaxial palea surface only and obtain percentage present/absent
guaduafloret_river_palea_bilobate_horizontal_per <- guaduafloret_river %>% 
  group_by(palea_bilobate_horizontal) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_river_palea_bilobate_horizontal_per)

#Create labeled pie chart of percentage of river specimens with simple papillae on palea abaxial surface
ggplot(guaduafloret_river_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
              absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


river_pie_palea_bilobate_horizontal <- ggplot(guaduafloret_river_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
  on palea abaxial surface")) +
  ggtitle("River") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("river_pie_palea_bilobate_horizontal.tiff", plot = river_pie_palea_bilobate_horizontal, width = 6, height = 4, units = "in", dpi = 600)
```


## Pie charts - specimens from forest species

##Filtering dataframe for forest species only
```{r}
#Filter dataframe of foliar micromorphology to include only forest species
guaduafloret_forest <- df_guaduafloret %>%
  filter(habitat == "forest")
```

##Forest: Pie chart of simple papillae on abaxial lemma surface presence/absence
```{r}
#Group by simple papillae on abaxial lemma surface only and obtain percentage present/absent
guaduafloret_forest_lemma_papillae_simple_per <- guaduafloret_forest %>% 
  group_by(lemma_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_forest_lemma_papillae_simple_per)

#Create labeled pie chart of percentage of forest specimens with simple papillae on lemma abaxial surface
ggplot(guaduafloret_forest_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on lemma abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


forest_pie_lemma_simple_papillae <- ggplot(guaduafloret_forest_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on lemma abaxial surface")) +
  ggtitle("Forest") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("forest_pie_lemma_simple_papillae.tiff", plot = forest_pie_lemma_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##Forest: Pie chart of simple papillae on abaxial palea surface presence/absence
```{r}
#Group by simple papillae on abaxial palea surface only and obtain percentage present/absent
guaduafloret_forest_palea_papillae_simple_per <- guaduafloret_forest %>% 
  group_by(palea_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_forest_palea_papillae_simple_per)

#Create labeled pie chart of percentage of forest specimens with simple papillae on palea abaxial surface
ggplot(guaduafloret_forest_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


forest_pie_palea_simple_papillae <- ggplot(guaduafloret_forest_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on palea abaxial surface")) +
  ggtitle("Forest") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("forest_pie_palea_simple_papillae.tiff", plot = forest_pie_palea_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##Forest: Pie chart of bilobate horizontally elongated silica bodies on abaxial palea surface presence/absence
```{r}
#Group by bilobate horizontally elongated silica bodies on abaxial palea surface only and obtain percentage present/absent
guaduafloret_forest_palea_bilobate_horizontal_per <- guaduafloret_forest %>% 
  group_by(palea_bilobate_horizontal) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_forest_palea_bilobate_horizontal_per)

#Create labeled pie chart of percentage of forest specimens with simple papillae on palea abaxial surface
ggplot(guaduafloret_forest_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
              absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


forest_pie_palea_bilobate_horizontal <- ggplot(guaduafloret_forest_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
  on palea abaxial surface")) +
  ggtitle("Forest") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("forest_pie_palea_bilobate_horizontal.tiff", plot = forest_pie_palea_bilobate_horizontal, width = 6, height = 4, units = "in", dpi = 600)
```


## Pie charts - specimens from savanna species

##Filtering dataframe for savanna species only
```{r}
#Filter dataframe of foliar micromorphology to include only forest and riparian species
guaduafloret_savanna <- df_guaduafloret %>%
  filter(general_habitat == "savanna")
```

##Savanna: Pie chart of simple papillae on abaxial lemma surface presence/absence
```{r}
#Group by simple papillae on abaxial lemma surface only and obtain percentage present/absent
guaduafloret_savanna_lemma_papillae_simple_per <- guaduafloret_savanna %>% 
  group_by(lemma_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_savanna_lemma_papillae_simple_per)

#Create labeled pie chart of percentage of savanna specimens with simple papillae on lemma abaxial surface
ggplot(guaduafloret_savanna_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on lemma abaxial surface")) +
  scale_fill_manual(labels = c("present", "absent"), values = c("#f7b54a", "#71caeb")) +
  theme_void()


savanna_lemma_simple_papillae <- ggplot(guaduafloret_savanna_lemma_papillae_simple_per, aes(x = "", y = perc, fill = lemma_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on lemma abaxial surface")) +
  ggtitle("Savanna") +
  scale_fill_manual(labels = c("present", "absent"), values = c("#f7b54a", "#71caeb")) +
  theme_void()

ggsave("savanna_pie_lemma_simple_papillae.tiff", plot = savanna_lemma_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##Savanna: Pie chart of simple papillae on abaxial palea surface presence/absence
```{r}
#Group by simple papillae on abaxial palea surface only and obtain percentage present/absent
guaduafloret_savanna_palea_papillae_simple_per <- guaduafloret_savanna %>% 
  group_by(palea_papillae_simple) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_savanna_palea_papillae_simple_per)

#Create labeled pie chart of percentage of general forest specimens with simple papillae on palea abaxial surface
ggplot(guaduafloret_savanna_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("present", "absent"), values = c("#f7b54a", "#71caeb")) +
  theme_void()

savanna_pie_palea_simple_papillae <- ggplot(guaduafloret_savanna_palea_papillae_simple_per, aes(x = "", y = perc, fill = palea_papillae_simple)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Simple papillae on palea abaxial surface")) +
  ggtitle("Savanna") +
  scale_fill_manual(labels = c("present", "absent"), values = c("#f7b54a", "#71caeb")) +
  theme_void()

ggsave("savanna_pie_palea_simple_papillae.tiff", plot = savanna_pie_palea_simple_papillae, width = 6, height = 4, units = "in", dpi = 600)
```

##Savanna: Pie chart of bilobate horizontally elongated silica bodies on abaxial palea surface presence/absence
```{r}
#Group by bilobate horizontally elongated silica bodies on abaxial palea surface only and obtain percentage present/absent
guaduafloret_savanna_palea_bilobate_horizontal_per <- guaduafloret_savanna %>% 
  group_by(palea_bilobate_horizontal) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

#view(guaduafloret_savanna_palea_bilobate_horizontal_per)

#Create labeled pie chart of percentage of savanna specimens with bilobate horizontally elongated silica bodies on palea abaxial surface
ggplot(guaduafloret_savanna_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
              absent or present on palea abaxial surface")) +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()


savanna_pie_palea_bilobate_horizontal <- ggplot(guaduafloret_savanna_palea_bilobate_horizontal_per, aes(x = "", y = perc, fill = palea_bilobate_horizontal)) +
  geom_col() +
  coord_polar(theta = "y") +
  guides(fill = guide_legend(title = "Bilobate horizontally elongated silica bodies
  on palea abaxial surface")) +
  ggtitle("Savanna") +
  scale_fill_manual(labels = c("absent", "present"), values = c("#71caeb", "#f7b54a")) +
  theme_void()

ggsave("savanna_pie_palea_bilobate_horizontal.tiff", plot = savanna_pie_palea_bilobate_horizontal, width = 6, height = 4, units = "in", dpi = 600)
```