---
title: "Guadua Leaf Micromorphology Analysis"
author: "Elizabeth McMurchie"
date: "2022-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load appropriate libraries, installing if needed
```{r}
library(RRPP) #factorial manova uses this 
library(geomorph) #factorial manova uses this
library(tidyverse) #general tidyverse packages
library(ade4) #needed for PCOA
library(vegan) #needed for PCoA
library(pander) #for model comparison
```

## Read in data as dataframe
```{r}
guadualeaf <- read_csv("../data/tidy/guadualeaf.csv", col_names = TRUE, na = "x")
```
## Drop columns with NA and no variation and create dataframe with columsn as factors
```{r}
guadualeaf <- guadualeaf %>%
  select(-c(ad_stomata_freq, ad_papillae_overarch, ad_triangular_sub_cells,
            ad_dome_sub_cells, ad_parallel_sub_cells, ab_ridged_saddle_intercostal,
            ad_ridged_saddle_intercostal)) # remove columns with NA or no variation

df_guadualeaf = data.frame(guadualeaf, stringsAsFactors = TRUE) # create dataframe

df_guadualeaf <- data.frame(lapply(df_guadualeaf,as.factor)) # set columns as factors

```

## Analysis


### PCoA 
The distance matrix for the binary data is obtained using simple matching coefficient. The PCoA from the distance matrix can be used to visualize the pairwise distances between the micromorphology of individual *Guadua* specimens.
```{r}
guadua_y_binary <- df_guadualeaf[,9:ncol(df_guadualeaf)] # reads in the binary data only 
guadua_y_binary_df <- data.frame(lapply(guadua_y_binary,function(x) as.numeric(levels(x))[x]))
guadua_y_dist <- dist.binary(guadua_y_binary_df, method = 2, diag = FALSE, upper = FALSE) # convert to distances using simple matching coefficient
guadua_y_dist_matrix <- as.matrix(guadua_y_dist) # create matrix using simple matching distances
PCoA <- cmdscale(guadua_y_dist_matrix, eig = TRUE, x.ret = TRUE, list. = TRUE) #from vegan - run PcoA
```

# Factorial MANOVA via RRPP excluding region - 2 groups for each
```{r}
leafdat <- rrpp.data.frame("guadua_y_binary_df" = guadua_y_binary_df,
                         "Habit" = as.factor(df_guadualeaf$general_habit),
                         "Habitat" = as.factor(df_guadualeaf$general_habitat)) 
model1.rrpp <- lm.rrpp(guadua_y_dist_matrix ~ leafdat$Habit + leafdat$Habitat, 
                       print.progress = FALSE)
anova(model1.rrpp)
```

# Factorial MANOVA via RRPP excluding region - 3 groups for each 
```{r}
leafdat2 <- rrpp.data.frame("guadua_y_binary_df" = guadua_y_binary_df,
                         "Habit" = as.factor(df_guadualeaf$habit),
                         "Habitat" = as.factor(df_guadualeaf$habitat)) 
model2.rrpp <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit + leafdat2$Habitat, 
                       print.progress = FALSE)
anova(model2.rrpp)

```

#Factorial MANOVA via RRPP with region (two groups for habit and habitat)
```{r}
leafdat3 <- rrpp.data.frame("guadua_y_binary_df" = guadua_y_binary_df,
                       "Habit" = as.factor(df_guadualeaf$general_habit),
                       "Habitat" = as.factor(df_guadualeaf$general_habitat),
                       "Region" = as.factor(df_guadualeaf$region)) 
model3.rrpp <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habit + leafdat3$Habitat + leafdat3$Region,  
                       print.progress = FALSE)
anova(model3.rrpp)

```

## Pairwise comparisons for region
```{r}
### Factorial model. Need your data matrix in the rdf
rdf <- rrpp.data.frame(morph.dist = guadua_y_dist, 
                       habit = df_guadualeaf$general_habit,
                       habitat = df_guadualeaf$general_habitat,
                       region = df_guadualeaf$region)
fit <- lm.rrpp(morph.dist ~ habit + habitat + region, data = rdf)

anova(fit)

reveal.model.designs(fit) # This step appears not to be working properly

null <- lm.rrpp(morph.dist ~ region, data = rdf)

PW <- pairwise(fit,fit.null = null,groups = rdf$region)

summary(PW,test.type = "dist") 

```


## Model comparison using likelihood ratio test (LTR)

Setup for model comparison:
```{r}
guadua_Habit <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habit, 
                 print.progress = FALSE)
guadua_Habitat <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habitat, 
                   print.progress = FALSE)
guadua_Region <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Region, 
                  print.progress = FALSE)
guadua_Habit_Region <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habit + leafdat3$Region, 
                        print.progress = FALSE)
guadua_Habitat_Region <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habitat + leafdat3$Region,
                          print.progress = FALSE)
guadua_Habit_Habitat <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habit + leafdat3$Habitat,
                         print.progress = FALSE)
guadua_full <- lm.rrpp(guadua_y_dist_matrix ~ leafdat3$Habit + leafdat3$Habitat + leafdat3$Region, 
                print.progress = FALSE)
```
#### RRPP MODEL COMPARISON 
For our model comparison, we used model.comparison() from the RRPP package using the log likelihood method. From this, we are able to determine that the full model (habit, habitat, and region) is the best fit based on it has the highest log-likelihood score and the lowest AIC score. 
```{r}
anova(guadua_full)

modelComp1 <- model.comparison(guadua_full, 
                             guadua_Habit_Habitat,
                             guadua_Habitat_Region, 
                             guadua_Habit_Region, 
                             guadua_Habitat, 
                             guadua_Habit, 
                             guadua_Region,
                             type = "logLik", tol = 0.01)
modelComp1.summ <- as.data.frame(summary(modelComp1))

pandoc.table(modelComp1.summ,
             style = "grid", 
             plain.ascii = TRUE)
```



