---
title: "Guadua Leaf Micromorphology Analysis"
author: "Elizabeth McMurchie, Devin Molnau, and Dean Adams"
date: "2025-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load appropriate libraries, installing if needed
```{r}
library(RRPP) #factorial manova uses this 
library(geomorph) #factorial manova uses this
library(tidyverse) #general tidyverse packages
library(ade4) #needed for PCOA
library(vegan) #needed for PCoA
library(pander) #for model comparison
```

## Read in data as dataframe
```{r}
guadualeaf <- read_csv("../data/tidy/guadualeaf.csv", col_names = TRUE, na = "x")
```

## Drop columns with NA and no variation and create dataframe with columns as factors
```{r}
guadualeaf <- guadualeaf %>%
  select(-c(ad_stomata_freq, ad_papillae_overarch, ad_triangular_sub_cells,
            ad_dome_sub_cells, ad_parallel_sub_cells, ab_ridged_saddle_intercostal,
            ad_ridged_saddle_intercostal)) # remove columns with NA or no variation

df_guadualeaf = data.frame(guadualeaf, stringsAsFactors = TRUE) # create dataframe

df_guadualeaf <- data.frame(lapply(df_guadualeaf,as.factor)) # set columns as factors

```

## Analysis

### PCoA 
The distance matrix for the binary data is obtained using simple matching coefficient. The PCoA from the distance matrix can be used to visualize the pairwise distances between the micromorphology of individual *Guadua* specimens.
```{r}
guadua_y_binary <- df_guadualeaf[,9:ncol(df_guadualeaf)] # reads in the binary data only 
guadua_y_binary_df <- data.frame(lapply(guadua_y_binary,function(x) as.numeric(levels(x))[x]))
guadua_y_dist <- dist.binary(guadua_y_binary_df, method = 2, diag = FALSE, upper = FALSE) # convert to distances using simple matching coefficient
guadua_y_dist_matrix <- as.matrix(guadua_y_dist) # create matrix using simple matching distances
PCoA <- cmdscale(guadua_y_dist_matrix, eig = TRUE, x.ret = TRUE, list. = TRUE) #from vegan - run PcoA
```

### Factorial MANOVA via RRPP - 3 groups for each 
```{r}
leafdat2 <- rrpp.data.frame("guadua_y_binary_df" = guadua_y_binary_df,
                         "Habit" = as.factor(df_guadualeaf$habit),
                         "Habitat" = as.factor(df_guadualeaf$habitat)) 
model2.rrpp <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit * leafdat2$Habitat, 
                       print.progress = FALSE)
anova(model2.rrpp)

```

### Pairwise comparison of habitats
```{r}

mod.habitat <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habitat,

                       print.progress = FALSE)

PW.habitat <- pairwise(mod.habitat, groups = df_guadualeaf$habitat)

summary(PW.habitat, test = "dist")
```

### Pairwise comparison of habits
```{r}
mod.habit <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit,

                       print.progress = FALSE)

PW.habit <- pairwise(mod.habit, groups = df_guadualeaf$habit)

summary(PW.habit, test = "dist")
```

### Exploration of interaction between habitat and habit
```{r}
gps.all <- paste(df_guadualeaf$habitat, df_guadualeaf$habit)

PW.all <- pairwise(model2.rrpp, groups = gps.all)

summary(PW.all, test = "dist") #test of means

summary(PW.all, test = "var")  #test of within-group variation
```


### Model comparison using likelihood ratio test (LTR)

Setup for model comparison:
```{r}
guadua_Habit2 <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit, 
                 print.progress = FALSE) # habit alone
guadua_Habitat2 <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habitat, 
                   print.progress = FALSE) # habitat alone
guadua_Habit_Habitat2 <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit + leafdat2$Habitat,
                         print.progress = FALSE) # considering habit and habitat with no interaction
guadua_full2 <- lm.rrpp(guadua_y_dist_matrix ~ leafdat2$Habit * leafdat2$Habitat, 
                print.progress = FALSE) # considering habitat and habitat with interaction

```

#### RRPP model comparison
For our model comparison, we used model.comparison() from the RRPP package using the log likelihood method. From this, we are able to determine that the model including habit and habitat is the best fit based on it has the highest log-likelihood score and the lowest AIC score. 
```{r}

modelComp2 <- model.comparison(guadua_full2,
                               guadua_Habit_Habitat2,
                               guadua_Habitat2, 
                               guadua_Habit2, 
                             type = "logLik", tol = 0.01)
modelComp2.summ <- as.data.frame(summary(modelComp2))

pandoc.table(modelComp2.summ,
             style = "grid", 
             plain.ascii = TRUE)
```

