---
title: "Guadua Floret Micomorphology Analysis"
author: "Elizabeth McMurchie"
date: "2022-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## load appropriate libraries, installing if needed
```{r}
library(RRPP) #factorial manova uses this 
library(geomorph) #factorial manova uses this
library(tidyverse) #general tidyverse packages
library(ade4) #needed for PCOA
library(vegan) #needed for PCoA
library(pander) #for model comparison
```

## Read in data as dataframe
```{r}
guaduafloret <- read_csv("../data/tidy/guaduafloret.csv", col_names = TRUE, na = "x")
```
## Drop columns with NA and no variation and create dataframe with columsn as factors
```{r}
guaduafloret <- guaduafloret %>%
  select(-c(lemma_stomata, palea_sulcus_prickle, palea_sulcus_macro, palea_stomata_freq, palea_subsidiary_triangular, palea_subsidiary_dome, palea_subsidiary_parallel)) # remove columns with NA or no variation

df_guaduafloret = data.frame(guaduafloret, stringsAsFactors = TRUE) # create dataframe

df_guaduafloret <- data.frame(lapply(df_guaduafloret,as.factor)) # set columns as factors

```

## Analysis

### Correlation
The correlation is necessary to complete the PCoA. Additionally, viewing the correlation dataframe is useful for understanding which traits are correlated.
```{r}
guadua_floret_y_binary <- df_guaduafloret[,10:ncol(df_guaduafloret)] # reads in the binary data only 
guadua_floret_y_binary_df <- data.frame(lapply(guadua_floret_y_binary,function(x) as.numeric(levels(x))[x]))
correlation_y_guadua_floret <- data.frame(cor(guadua_floret_y_binary_df)) #calculated the correlation of the Y values
```

### PCoA 
The distance matrix for the binary data is obtained using simple matching coefficient. The PCoA from the distance matrix can be used to visualize the pairwise distances between the micromorphology of individual *Guadua* specimens.
```{r}
guadua_floret_y_dist <- dist.binary(guadua_floret_y_binary_df, method = 2, diag = FALSE, upper = FALSE) # convert to distances using simple matching coefficient
guadua_floret_y_dist_matrix <- as.matrix(guadua_floret_y_dist) # create matrix using simple matching distances
floret_PCoA <- cmdscale(guadua_floret_y_dist_matrix, eig = TRUE, x.ret = TRUE, list. = TRUE) #from vegan - run PcoA
```

# Factorial MANOVA via RRPP excluding region - 2 groups for each
```{r}
floretdat <- rrpp.data.frame("guadua_floret_y_binary_df" = guadua_floret_y_binary_df,
                         "Habit" = as.factor(df_guaduafloret$general_habit),
                         "Habitat" = as.factor(df_guaduafloret$general_habitat)) 
model1.rrpp <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat$Habit + floretdat$Habitat, 
                       print.progress = FALSE)
anova(model1.rrpp)
```

# Factorial MANOVA via RRPP excluding region - 3 groups for each 
```{r}
floretdat2 <- rrpp.data.frame("guadua_floret_y_binary_df" = guadua_floret_y_binary_df,
                         "Habit" = as.factor(df_guaduafloret$habit),
                         "Habitat" = as.factor(df_guaduafloret$habitat)) 
model2.rrpp <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat2$Habit + floretdat2$Habitat, 
                       print.progress = FALSE)
anova(model2.rrpp)

```

#Factorial MANOVA via RRPP with region (two groups for habit and habitat)
```{r}
floretdat3 <- rrpp.data.frame("guadua_floret_y_binary_df" = guadua_floret_y_binary_df,
                       "Habit" = as.factor(df_guaduafloret$general_habit),
                       "Habitat" = as.factor(df_guaduafloret$general_habitat),
                       "Region" = as.factor(df_guaduafloret$region)) 
model3.rrpp <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habit + floretdat3$Habitat + floretdat3$Region,  
                       print.progress = FALSE)
anova(model3.rrpp)

```

## Model comparison using likelihood ratio test (LTR)

Setup for model comparison:
```{r}
guadua_Habit <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habit, 
                 print.progress = FALSE)
guadua_Habitat <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habitat, 
                   print.progress = FALSE)
guadua_Region <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Region, 
                  print.progress = FALSE)
guadua_Habit_Region <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habit + floretdat3$Region, 
                        print.progress = FALSE)
guadua_Habitat_Region <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habitat + floretdat3$Region,
                          print.progress = FALSE)
guadua_Habit_Habitat <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habit + floretdat3$Habitat,
                         print.progress = FALSE)
guadua_full <- lm.rrpp(guadua_floret_y_dist_matrix ~ floretdat3$Habit + floretdat3$Habitat + floretdat3$Region, 
                print.progress = FALSE)
```

#### RRPP MODEL COMPARISON 
For our model comparison, we used model.comparison() from the RRPP package using the log likelihood method. From this, we are able to determine that the Y.group2 model is the best fit based on it has the highest log-likelihood score and the lowest AIC score. 
```{r}
# Including region
anova(guadua_full)

modelComp1 <- model.comparison(guadua_full, 
                             guadua_Habit_Habitat,
                             guadua_Habitat_Region, 
                             guadua_Habit_Region, 
                             guadua_Habitat, 
                             guadua_Habit, 
                             guadua_Region,
                             type = "logLik", tol = 0.01)
modelComp1.summ <- as.data.frame(summary(modelComp1))

pandoc.table(modelComp1.summ,
             style = "grid", 
             plain.ascii = TRUE)


# This one excluding region, which wasn't significant 

anova(guadua_Habit_Habitat)

modelComp2 <- model.comparison(guadua_Habit_Habitat,
                             guadua_Habitat, 
                             guadua_Habit, 
                             type = "logLik", tol = 0.01)
modelComp2.summ <- as.data.frame(summary(modelComp2))

pandoc.table(modelComp2.summ,
             style = "grid", 
             plain.ascii = TRUE)
```